---
title: "For Loop Challenges"
author: "Aaron Honig"
date: "October 20, 2015"
output: html_document
---

For Loops Challenges:

1.   *Print each value of this matrix.*

        So here, I just told for loop to print each value of the matrix. 

```{r} 
      mat <- matrix(1:100, nrow = 10) # Enter matrix values, number of rows
      for(i in 1:100){ # Tell for loop over which values to integrate
        print(mat[i]) # Print each value of i for the constructed matrix
      }
``` 

2.  *Multiply each value in this matrix by 7 and store it in a 10 x 10 matrix.*

        So here, I created a matrix with ten rows and columns, and within the for loop, multiplied each row (i) and column (j) by seven.  

```{r}
      mat <- matrix(1:100, nrow = 10) # Enter matrix values, number of rows
      for(i in 1:10){ # Tell for loop over which rows to integrate
         for(j in 1:10){ # Tell for loop over which columns to integrate
            mat[i,j] = mat[i,j]*7 # Multiple each row/column value by 7
            
      } 
      } 
      mat  # Print new matrix outside of for loop
```

3.   *a) Print these values as part of a string that looks something like                'n = 16'.*
     *b) Now modify this loop to store these strings in a new vector called               counts.*
 
        So here, I rounded 15 numbers randomly drawn from a uniform distribution between 10 and 100.  For each value in that vector, I pasted "n = " before that value, and placed it into a new vector called "counts".
        
```{r}
      set.seed(1) # Enter seed number to keep random values constant
      x <- round(runif(min = 10, max = 100, n = 15)) 
            # Put 15 random numbers between 10 and 100, drawn from a uniform                        distribution, then rounded them to the nearest hold number into vector "x"
      counts <- vector(length = 15) # Create blank vector of length 15
      for(i in 1:length(counts)){ # Tell for loop to integrate length of vector
        counts[i] <- paste("n = ", x[i])                                                    # For each value of vector, paste together "n = " and the ordinal value of that           value
      }
      counts $ # Print count vector
```      
      
4.    *Make a vector for which each entry is 2 raised to the power of it’s index (ex: the 3rd item in the vector is equal to 2^3).*

      So here, I created a new vector of length ten, and raised two to the power of each ordinal value of that vector. 

```{r}
      power <- vector(length = 10) # Create blank vector of length 10
      for(i in 1:10){ # Tell for loop to integrate over length of vector
        power[i] <- 2^i  # For each value of vector, calculate 2 to the power of that                             value
      }
      power # Print power vector
```

5.    *Make a matrix where each entry, using indexes i for row and j for column, is equivalent to i x j. Your final output should look like:*
  
      So here, I created a blank matrix of 12 rows and columns. For each row (i) and column (j) value, I multiplied each ordinal row value by each ordinal column value.  

```{r}
      mat <- matrix(ncol = 12, nrow = 12) # Create blank matrix of 12 columns, rows
      for(i in 1:12){ # Tell for loop over which rows to integrate 
        for(j in 1:12){ # Tell for loop over which columns to integrate
          mat[i,j] <- i*j # Multiply each matrix value by ordinal row and column
      }
      }
      mat # Print calculated matrix
```

6.    *Make a vector where each entry is TRUE or FALSE, based on whether it’s index is even or odd.*

      So here, I tried two different things.  First, I created a vector of length ten, loaded a package called gtools, created a blank vector, and used the function "even" for each ordinal value of the vector to return a logical value for evenness.  Secondarily, I created a vector of length ten, and wrote an if/then statement returning an affirmative phrase if exactly half of the value was equal to half of that value rounded to the nearest whole number (only true if the value is even).  If it were not true, a negative phrase is returned. 

```{r}
      x <- 1:10 # Create vector of integers 1 to 10
      library(gtools) # Load gtools package, which includes "even" function
      logic <- vector(length = 10) # Create blank vector of length 10
      for(i in 1:10){ # Tell for loop over which values to integrate
        logic[i] <- even(x[i]) # For each vector value, print logical value for                                        evenness
      }
      logic # Print logic vector

      x <- c(1:10) # Create vector of integers 1 to 10
      for(i in 1:10){ # Tell for loop over which values to integrate
        if(x[i]/2 == round(x[i]/2)){ # Write conditional if then to determine if                 exactly one half of vector value equals the rounded integer of one half of             vector value (only true if vector is divisible by 2; or even)
        print(c(x[i],"Hell Ya, Man!")) # If True, then print in the affirmative
      } else { 
        print(c(x[i],"No Way Dude!")) # If not true, then print in the negative
      }
      }
```

  7.  *Using dplyr, calculate the mean abundance of each taxonomic group, what do you notice about the output? Is it what you would expect?*
  
      So here, I created a character vector called taxa, randomly sampled that vector 100 times with replacement, generated 100 numbers randomly generated from a uniform distribution between 10 and 500, joined both in a single dataframe, capitalized the dataframe and columns names, and calculated the mean abundance values for each taxa. 
    
```{r}
      taxa <- c('Coral', 'fish', 'Fish', 'Phytoplankton', 'coral',                      'phytoplankton', 'zooplankton', 'Zooplankton', 'Echinoderms',                   'echinoderms', 'Cephalopods', 'cephalopods') # Inform vector "taxa"
      taxa_values <- sample(taxa, size = 100, replace = TRUE) # Randomly sample 100            values from taxa vector with replacement
      set.seed(1) # Enter seed number to keep random values constant
      counts <- round(runif(min = 10, max = 500, n = 100)) # Put 15 random numbers             between 10 and 100, drawn from a uniform distribution, then rounded them to the         nearest hold number into vector "counts"
      taxa_counts <- data.frame(taxa = taxa_values, abundance = counts) # Create               dataframe "taxa_counts" with taxa_values and count vectors
      colnames(taxa_counts) <- c("Taxa", "Abundance") # Rename columns
      TaxaCounts <- taxa_counts # Rename dataset
      ddply(TaxaCounts, ~Taxa,summarise, mean=mean(Abundance)) # Summarize mean                Abundance values, grouped by Taxa
```
   
   So no, there are duplicate names because some are not capitalized.  So I need to standardize the capitalization of the taxa names. So here I use a for loop to take the first letter of every taxa name, make that letter capitalized, paste it to the rest of the taxa name, and find the mean abundance for each of the now-capitalized taxa names.
   
   
```{r}      
      for(i in 1:100){ # Tell for loop over which values to integrate
      TaxaCounts[i,1] <- paste(toupper(substring(TaxaCounts[i,1], 1, 1)),               substring(TaxaCounts[i,1], 2,15), sep = "", collapse = " ")
      } # Take first letter of each name in the first column of TaxaCounts using               "substring", make that letter uppercase with "toupper" and paste it to                 the next 15 letters of each name, getting rid of extra spaces with                     "collapse"
      ddply(TaxaCounts, ~Taxa,summarise,mean=mean(Abundance)) # Summarize mean                 Abundance values, grouped by Taxa
```


