---
title: "Hmwk2"
author: "Aaron Honig"
date: "September 25, 2015"
output: html_document
---

Homework #2


1. To make a scatterplot of life expectancy as a function of gdp per capita, assigning each continent an independent color and shape, I assigned a color to continents in the ggplot aes, and shape to aes within geom_point. 

```{r}
library(ggplot2)
library(gapminder)
library(plyr)
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + geom_point(aes(shape = continent))
```


2.  To make the same figure with a log transformed x-axis, I added scale_x_log10(). 

```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + geom_point(aes(shape = continent)) + scale_x_log10()
```

3. To add a smoothed linear regression line over the plot, I added geom_smooth(method = lm), and moved the color assignment to geom_point so that geom_smooth was not performed for each continent independently. 

```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point(aes(shape = continent, color = continent)) + scale_x_log10() + geom_smooth(method = lm)
```

4.  To facet wrap density plots of life expectancy for each continent by year, I used "fill", informed facet wrap by year, assigned 4 columns, and reduced the opacity by 0.5. 

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) + facet_wrap( ~ year, ncol=4) + geom_density(alpha = 0.5)
```


5.  To facet wrap boxplots of life expectancy for each continent by year, I assigned geom_boxplot, and used color with the ggplot aes. I also rotated the x-axis labels to make them more legible. 

```{r}
ggplot(data = gapminder, aes(x = continent, y = lifeExp, color = continent)) + facet_wrap( ~ year, ncol=4) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45)) + labs(x = "incontinent")
```

6.  To plot overall life expectancy by continent, I used "fill" within the ggplot aes and reduced opacity within geom_density by 0.5. 

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) + geom_density(alpha = 0.5)
```

7.  To do the same thing with only Asia, I subset the data within ggplot using a pipe, and filled in with green using scale_fill_manual values.  I added a vertical line, calculated as the mean life expectancy of the subset Asia dataset, and retitled the figure. 

```{r}
ggplot(data = subset(gapminder, continent %in% c("Asia")), aes(x = lifeExp, fill = continent)) + geom_density(alpha = 0.5) + scale_fill_manual(values = c("green1")) + geom_vline(data = subset(gapminder, continent %in% c("Asia")), aes(xintercept = mean(lifeExp))) + guides(fill=FALSE) + ggtitle("Life expectancy in Asia")
```

8. To do the same thing, only facet wrapped with all continents, I used fill within the ggplot aes, facet wrapped by continent with 3 columns, and used ddply within geom_vline to draw mean life expectancy values within each facet wrapped figure. 

```{r}
ggplot(data = gapminder, aes(x = lifeExp, fill = continent))  + geom_density(alpha = 0.7) + geom_vline(data = ddply(gapminder, "continent", summarize, mean = mean(lifeExp)), aes(xintercept = mean)) + facet_wrap( ~ continent, ncol=3)
```

9. a. Here I uploaded the data as a txt file.

```{r}
Gapminder_New <- read.csv("~/Desktop/Gapminder_New.txt")
```

   b. The life expectancy column name was misspelled. 
   
```{r}
mean_lifeExp <- mean(Gapminder_New$lifeExp)
mean_lifeExp
```
   
   c. The column names were not properly assigned as a vector using c().
   
```{r}
small_set <- Gapminder_New[c(1, 2, 3, 4, 1300:1304), c('country', 'continent', 'year')]
small_set
```
      
   d. NA values within the gdpPercap column needed to be ignored using na.rm.
   
```{r}
mean_gdp <- mean(Gapminder_New$gdpPercap, na.rm = TRUE)
mean_gdp
```      
  
  e.  "Which" is for losers.  Here I used filter to find the data for the country with max life expectancy, then used bracket to spit out the country name.  Throw me a real challenge next time.....
   
```{r}
max_country <- filter(Gapminder_New, lifeExp == max(lifeExp))["country"]
max_country
```
